---
title: 'Simulation - Part 1'
author: M. L. DeBusk-Lane
date: '2021-01-22'
slug: []
categories: [Simulation]
tags: [multi-level regression]
subtitle: ''
summary: 'Linear and Linear Multilevel Data Simulation'
authors: []
lastmod: '2021-01-22T06:57:57-05:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
output: 
  html_document:
    theme: journal
    highlight: zenburn
---
 
There is something very satisfying, reassuring, and down-right compelling to making your own data, establishing known paramters, and then seeing how particular methods work to recover them. In graduate school I found this utility increasingly useful in exploring methods and, more importantly, learning them. 

## Linear Data Generation 

https://aosmith.rbind.io/2018/01/09/simulate-simulate-part1/

$$y_t = \beta_0 + \beta_1*I_{(group_t=\textit{group2})} + \epsilon_t$$


+ $y_t $ is the observed values for the quantitative response variable; $t$ goes from 1 to the number of observations in the dataset. Said aother way, this is the outcome, response variable, or what we are trying to predict with the right hand side of the equation.   
+ $\beta_0$ is the mean response variable when the group is *group1*. Intercept... where x = 0. 
+ $\beta_1$ is the difference in mean response between the two groups, *group2* minus *group1*. If and only if the grouping variable is dichotomous or binary. 
+ The indicator variable, $I_{(group_t=\textit{group2})}$, is 1 when the group is *group2* and 0 otherwise, so $\beta_1$ only affects the response variable for observations in *group2*. This is why you can look at the intercept to computer the mean of the group that = 0. 
+ $\epsilon_t$ is the random variation present for each observation that is not explained by the group variable.  These are assumed to come from an iid normal distribution with a mean of 0 and some shared variance, $\sigma^2$: $\epsilon_t \thicksim N(0, \sigma^2)$ This is often called error. 

For reproducibility, lets grab a seed. Quick fact: If you dont explicitly note a new `set.seed(interger here)` the seed is created from the current time and process ID from your local? There's actually a ton of information if you take a look at the help file for `set.seed()` You're welcome. 

```{r}
set.seed(23120)
```


Let's grab some packages/libraries. I use pacman... because just listing a bunch of libraries and hoping you have them installed is pretty rude. Just run the code, it'll download what you need and not what you don't. 

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  tidymodels
)
```
```{r}

```





 
## Multilevel Data Generation 
